A Journey from Wild to Textbook Data to Reproducibly Refresh the Wages Data from the National Longitudinal Survey of Youth Database

### 2022-02-17 paper/paper.Rmd
updated the takeaways

Reading the section titled ["Topical Guide to the Data"](https://www.nlsinfo.org/content/cohorts/nlsy79/topical-guide/employment/work-experience) in the guide, suggests that it should be calculated based on the variable "number of weeks worked since last interview". This would remove periods of unemployment which makes sense when measuring experience while actually working. Since it is only measured since the last interview, this needs to be cumulated for each survey year. This produced results more similar to the original data (see Subsection \@ref(takeaways)).

@@

There are two aspects of the original data that make the replication with exactly the same result is not obtained. The first is that the calculation of the work experience is not clearly articulated. @SingerJudithD2003Alda mention that the temporal variable is the years of experience since the first day of work. However, this exact same variable is not explicitly available in the database. From the NLSY79 topical guide [@nlsy79guide], we find that several variables are tagged as work experience-related variables. One of them is the weeks of worked since the last interview. This is used to calculate the variable. It produces similar results, as shown in Figure \ref{fig:compare-xp-sw}.

Secondly, in the original dataset, wages were inflation-adjusted to 1990 prices. This is not done for the refreshed data because we plan to keep refreshing it as the data added to and released from the survey. Instead we have provided a function in the R package, `yowie`, for users to conduct the inflation adjustment when they are ready to analyse the data.

>>>>

Reading the section titled ["Topical Guide to the Data"](https://www.nlsinfo.org/content/cohorts/nlsy79/topical-guide/employment/work-experience) in the guide, suggests that it should be calculated based on the variable "number of weeks worked since last interview". This would remove periods of unemployment which makes sense when measuring experience while actually working. Since it is only measured since the last interview, this needs to be cumulated for each survey year. This produced results more similar to the original data (see Section \@ref(takeaways)).

@@

There are two aspects of the original data that we cannot replicate. The first is that the calculation of the work experience is not clearly articulated. @SingerJudithD2003Alda mention that the temporal variable is the years of experience since the first day of work. However, this exact same variable is not explicitly available in the database. From the NLSY79 topical guide [@nlsy79guide], we find that several variables are tagged as work experience-related variables. One of them is the weeks of worked since the last interview. This is used to calculate the variable. It produces a highly correlated results as shown in Figure \ref{fig:compare-xp-sw}.

Secondly, in the original dataset, wages were inflation-adjusted to 1990 prices. This is not done for the refreshed data because we plan to keep refreshing it as the data added to and released from the survey. Instead we have provided a function in the R package, `r yowie`, for users to conduct the inflation adjustment when they are ready to analyse the data.
